What's hot and what's not:\\Windowed developer topic analysis

### 2009-06-18 lda-paper.tex
- [X] Add an LDA example (Dogs/cats - but with developers)     - [X] Make it clear what we mean when we do "local topic analysis"           "Our analysis shows that optimizer was important but it had           been obscured, but would have been noticed using the more           local topic analysis." What does local topic mean? A smaller           window?           - In the introduction we try to clarify this and in MaxDB 7.500     - [X] Why were certain values chosen? • Sometimes some choices are not so clear/well explained: one month as          the length of time windows, 20 topics, 10 words

\newcommand{\affaddr}[1]{#1}
\newcommand{\aemail}[1]{#1}
%\numberofauthors{5}
\author{
%\alignauthor
Abram Hindle\\
%\affaddr{David Cheriton School of Computer Science}\\
\affaddr{University of Waterloo}\\
\affaddr{Waterloo, Ontario}\\
\affaddr{Canada}\\
\aemail{ahindle@cs.uwaterloo.ca}
%\alignauthor
\and
Michael W. Godfrey\\
%\affaddr{David Cheriton School of Computer Science}\\
\affaddr{University of Waterloo}\\
\affaddr{Waterloo, Ontario}\\
\affaddr{Canada}\\
\aemail{migod@cs.uwaterloo.ca}
%\alignauthor
\and
Richard C. Holt\\
%\affaddr{David Cheriton School of Computer Science}\\
\affaddr{University of Waterloo}\\
\affaddr{Waterloo, Ontario}\\
\affaddr{Canada}\\
\aemail{holt@cs.uwaterloo.ca}
%\alignauthor
%\and
%Gregorio Robles\\
%\affaddr{GSyc}\\
%\affaddr{Universidad Rey Juan Carlos}\\
%\affaddr{Madrid}\\
%\affaddr{Spain}\\
%\aemail{grex@gsyc.escet.urjc.es}
}

@@

  As development on a software project progresses, developers will
  naturally shift their focus between different topics and tasks many
  times.  Managers and newcomer developers often seek ways of
  understanding what tasks have recently been worked on and how much
  effort has gone into each; for example, a manager might wonder what
  unexpected tasks occupied their team's attention during a period
  when they were supposed to have been implementing a set of new
  features.  Tools such as Latent Dirichlet Allocation (LDA) and
  Latent Semantic Indexing (LSI) can be used to analyze check-in
  comments over the entire lifetime of a project. Previous work on
  developer topic analysis has leveraged these tools to associate
  commit log comments with independent topics extracted from these
  commit log comments.  In this work, we use LDA  to analyze
  periods, such as months, within a project's lifetime to create a
  time-windowed model of changing development topics.  We propose 
  visualizations of this model that allows us to explore the evolving
  stream of topics of development occurring over time.  We demonstrate
  that windowed topic analysis offers advantages over topic analysis
  applied to a projects lifetime because many topics are quite local.

@@

We want to know about the topics of development that programmers
encountered while developing a software project. We want to know given
a certain period of time, what were the popular topics being worked on
during that period. We also want to know if any of those topics
occurred again, whether before or after our current period.  These
topics could help stakeholders determine what their coworkers were
working on, what were common issues, what requirements were being
worked on, etc.

@@

@@

In our first exploratory pass we wanted to see if LDA could provide
interesting topics extracted from a real system. We took the
repository of MySQL 3.23, extracted the commits, grouped them by 30
day non-overlapping windows, and then applied LDA to each of these
windows. We asked LDA to make 20 topics per window and we then
examined the top 10 most frequent words in that topic.  We found there
were common words across topic clusters, such as \emph{diffs},
\emph{annotate} and \emph{history}; since these words occur so often
in the MySQL 3.23 messages, we probably should be viewed as stop
words. There were notable transitional topics such as in the first
window the word \emph{BitKeeper} appears because MySQL adopted
Bitkeeper for their source control system yet in the following windows
there were no mentions of BitKeeper. \emph{RENAME} also only appeared
once in the first window and never again. Per each topic we tried to
identify the purpose of the topic; to our surprise we found that even
with only a little familiarity with the code base that naming the
topic was straightforward.

@@

2001 &  Aug &      TABLES row version \\
\hline         
2001 &  Sep &      update checksum merge \\
% 2001 &  Oct &      fixed fix \\
% 2001 &  Dec &      HPUX SCO fix \\
% \hline         
% 2002 &  Feb &      net buffer length  max\_allowed\_packet \\
% 2002 &  Mar &      small buf fix  \\
% \hline         
% 2002 &  May &      [popular] fix SCO OSF1 table\_name \\
% 2002 &  Nov &      HPUX11 compiler HP \\
% \hline         
% 2003 &  Feb &      Linux errno   \\
% 2003 &  Mar &      alarm bookmark bug \\
% \hline         
% 2003 &  Sep &      Auto logging merge windows distribution fix 64-bit 4.0 Cleanup \\
\hline
\end{tabular}
\caption{Sampled topics from MySQL 3.23, some with continuous topics. These tokens were pulled from LDA topic. Each token is a summary of one LDA generated topic from MySQL 3.23 commit comments.}
\label{tab:portability}
\end{table}

@@

Given all messages, we group the messages into windows. We could use
overlapping windows, but in this paper we use non-overlapping windows
of a set time unit because it simplifies analysis.  Windowing by time
allows for many levels of granularity.  We used one month as the
length of our time windows. While we could use different window
lengths for this study we think that a month is a sizable unit of
development, which is large enough to show shifts in focus, but coarse
enough not to show too much detail.

@@

Once we have our data windowed, we apply our Topic Analysis tool to
each window and extract the top $N$ topics. We used $20$ topics in our
case studies. Our Topic Analysis tool is an implementation of LDA, but
we could have used other similar tools like LSI, but previous work
showed more promising topic analysis results with LDA.  We note that
this step is a slow one, as executing even one instance of LDA
involves significant computation, and we perform LDA once per window.
Figure \ref{fig:commits} shows how LDA is applied to a set of
messages, and how the topics are extracted and related to the
messages.

@@

Our input for topic similarity is the top 10 most common words of each
topic.  Each topic is compared to each other topic in the system,
given a certain threshold of top 10 word similarity, such as 8 out of
top 10 matching words. We then apply the transitive closure on similar
topics to this topic similarity matrix; this is similar to modelling
topics as nodes and similarity as arcs, then fill flooding along the
similarity arcs until we have partitioned the topics into clusters of
similar topics. Figure \ref{fig:closure} illustrates clustering of
topics by topic similarity. These clusters of topics are called
trends. A trend is probably interesting if it contains more than one
topic.

@@

When we examined PostgreSQL, we did not find many trends with two or
more topics, using a similarity of $7/10$. Those trends that we did find
were not very large, lasting only 3 months at most. The first and second
largest trends were dedicated to the work of two external developers:
Dal Zotto, Dan McGuirk. The fifth largest trend related to work by
PostgreSQL developer D'Arcy. Other topics of the larger trends were
changes to the TODO, and time string formating topics
relating to timezones.

@@

One avenue of future work is automatic topic labelling. Given a word
distribution we should be able to automatically select a word or term
that describes that distribution. Potential external sources of topic
names include: software engineering taxonomies, ontologies and
standards.

>>>>

\newcommand{\affaddr}[1]{#1}
\newcommand{\aemail}[1]{#1}
%\numberofauthors{5}
\author{
%\alignauthor
Abram Hindle and Michael W. Godfrey and Richard C. Holt\\
%\affaddr{David Cheriton School of Computer Science}\\
\affaddr{University of Waterloo}\\
\affaddr{Waterloo, Ontario}\\
\affaddr{Canada}\\
\aemail{\{ahindle,migod,holt\}@uwaterloo.ca}
%\alignauthor
% \and
% Michael W. Godfrey\\
% %\affaddr{David Cheriton School of Computer Science}\\
% \affaddr{University of Waterloo}\\
% \affaddr{Waterloo, Ontario}\\
% \affaddr{Canada}\\
% \aemail{migod@cs.uwaterloo.ca}
% %\alignauthor
% \and
% Richard C. Holt\\
% %\affaddr{David Cheriton School of Computer Science}\\
% \affaddr{University of Waterloo}\\
% \affaddr{Waterloo, Ontario}\\
% \affaddr{Canada}\\
% ;\aemail{holt@cs.uwaterloo.ca}
%\alignauthor
%\and
%Gregorio Robles\\
%\affaddr{GSyc}\\
%\affaddr{Universidad Rey Juan Carlos}\\
%\affaddr{Madrid}\\
%\affaddr{Spain}\\
%\aemail{grex@gsyc.escet.urjc.es}
}

@@

  As development on a software project progresses, developers will
  naturally shift their focus between different topics and tasks many
  times.  Managers and newcomer developers often seek ways of
  understanding what tasks have recently been worked on and how much
  effort has gone into each; for example, a manager might wonder what
  unexpected tasks occupied their team's attention during a period
  when they were supposed to have been implementing a set of new
  features.  Tools such as Latent Dirichlet Allocation (LDA) and
  Latent Semantic Indexing (LSI) can be used to analyze commit log
  comments over the entire lifetime of a project. Previous work on
  developer topic analysis has leveraged these tools to associate
  commit log comments with independent topics extracted from these
  commit log comments.  In this work, we use LDA  to analyze
  periods, such as months, within a project's lifetime to create a
  time-windowed model of changing development topics.  We propose 
  visualizations of this model that allows us to explore the evolving
  stream of topics of development occurring over time.  We demonstrate
  that windowed topic analysis offers advantages over topic analysis
  applied to a projects lifetime because many topics are quite local.

@@

What happened in the previous maintenance or development iteration?
What were the developers working on? What was the developer in the
next cubicle working on? Which common topics and issues were dealt
with in the current release of the software? What were the topics of
the previous release? Given the requirements agreed to at the start
of the iteration did our developers work on them?  What else did they
work on?

We want to know about the topics of development that programmers
encountered while developing a software project. We want to know given
a certain period of time, what were the popular topics being worked on
during that period. We also want to know if any of those topics
occurred again, whether before or after our current period.  These
topics could help stakeholders determine what their coworkers were
working on, what were common issues, what requirements were being
worked on, etc. In order to extract and evaluate these topics we plan
to rely on topic analysis.

@@

% An example use of LDA
% * take a bunch of commit messages
% ** occurred during a period with bug fixing, adding a GUI and documentation changes
% ** LDA would find that bug fixing topic was associated to the words bug, fix, failure, ticket
% *** GUI - widget, panel, button, menu, window
% *** Documentation - chapter, section, figure and words shared between GUI and bug fixing
% * a particular commit would be associated with some of these topics 
% ** a commit about addressing a fix to a bug in a ticket would be heavily associated with the bug topic
% ** a commit that deals with documenting the GUI associated with GUI and Documentation
% ** a commit that was independant of all three could've induced a seperate topic altogether

Imagine that we are analyzing one month of development in a project
where the main focus was on bug fixing, adding a new GUI and
documentation updates relating to the system and the GUI. To use LDA
we'd get the word distributions of each commit messages, pass them to
LDA and tell LDA to find some topics, in this case we said 4. The 4
topics it discovered would probably be bug fixing, GUI, documentation
and other. These topics are essentially word distributions, for
example the bug fixing topic would include words like \emph{bug},
\emph{fix}, \emph{failure} and \emph{ticket}. The GUI topic would
include words like \emph{widget}, \emph{panel}, \emph{window}, and
\emph{menu}. The documentation topic would include words like
\emph{section}, \emph{figure}, and \emph{chapter}. The other topic
would include words found in topics generally independant of the other
topics. If we analyzed a commit of a bug fix it'd probably contain
words like \emph{ticket} and thus heavily associate with the bug fix
topic. A commit that was about user documentation regarding the GUI
would be associated with the documentation topic and the GUI topic
because it shared words from both topics. Where as a topic dealing
neither bugs, GUIs or documentation might be associated with the other
topic. This example demonstrates how LDA can associate 1 document (a
commit message) with more than one topic, how topics are related to
documents via their word distributions. LDA also would discover these
independant topics in an unsupervised manner, all the end user has to
do is name the topic based upon the most frequent words in that topic,
which is a word distribution.

@@

In our first exploratory pass we wanted to see if LDA could provide
interesting topics extracted from a real system. We took the
repository of MySQL 3.23, extracted the commits, grouped them by 30
day non-overlapping windows, and then applied LDA to each of these
windows. We asked LDA to make 20 topics per window and we then
examined the top 10 most frequent words in that topic.  We chose one
month because it was smaller than the time between minor releases but
large enough for there to be many commits to analyze. We chose 20
topics because past experimentation showed that few topics might
aggregate multiple unique topics while anymore topics seemed to dilute
the results and create some indistinct topics. 10 most frequent words
were chosen as sometimes 1 common word dominated topics so we wanted
to have enough words to distingush topics.

We found there
were common words across topic clusters, such as \emph{diffs},
\emph{annotate} and \emph{history}; since these words occur so often
in the MySQL 3.23 messages, we probably should be viewed as stop
words. There were notable transitional topics such as in the first
window the word \emph{BitKeeper} appears because MySQL adopted
Bitkeeper for their source control system yet in the following windows
there were no mentions of BitKeeper. \emph{RENAME} also only appeared
once in the first window and never again. Per each topic we tried to
identify the purpose of the topic; to our surprise we found that even
with only a little familiarity with the code base that naming the
topic was straightforward.

@@

2001 &  Aug &      TABLES row version \\
\hline         
2001 &  Sep &      update checksum merge \\
% 2001 &  Oct &      fixed fix \\
% 2001 &  Dec &      HPUX SCO fix \\
% \hline         
% 2002 &  Feb &      net buffer length  max\_allowed\_packet \\
% 2002 &  Mar &      small buf fix  \\
% \hline         
% 2002 &  May &      [popular] fix SCO OSF1 table\_name \\
% 2002 &  Nov &      HPUX11 compiler HP \\
% \hline         
% 2003 &  Feb &      Linux errno   \\
% 2003 &  Mar &      alarm bookmark bug \\
% \hline         
% 2003 &  Sep &      Auto logging merge windows distribution fix 64-bit 4.0 Cleanup \\
\hline
\end{tabular}
\caption{Sampled topics from MySQL 3.23, some with continuous topics. These tokens were pulled from the top 10 most common words found in LDA extracted topics. Each token is a summary of one LDA generated topic from MySQL 3.23 commit comments.}
\label{tab:portability}
\end{table}

@@

Given all messages, we group the messages into windows. We could use
overlapping windows, but in this paper we use non-overlapping windows
of a set time unit because it simplifies analysis.  Windowing by time
allows for many levels of granularity.  We used one month as the
length of our time windows. While we could use different window
lengths for this study we think that a month is a sizable unit of
development, which is large enough to show shifts in focus, but coarse
enough not to show too much detail. Choosing a month as the window
provided us with enough documents to analyze. One could overlap the
windows, which would increase the likelihood of trends, but for this
study we lacked the space and were more concerned if topics ever
repeated, overlapping might skew that result.

@@

Once we have our data windowed, we apply our Topic Analysis tool to
each window and extract the top $N$ topics. We used $20$ topics in our
case studies because we experimented before and that $20$ was about
all we can handle before too many topics seemed too indistinct from
each other to differeniate.

 Our Topic Analysis tool is an implementation of LDA, but
we could have used other similar tools like LSI, but previous work
showed more promising topic analysis results with LDA.  We note that
this step is a slow one, as executing even one instance of LDA
involves significant computation, and we perform LDA once per window.
Figure \ref{fig:commits} shows how LDA is applied to a set of
messages, and how the topics are extracted and related to the
messages.

@@

Our input for topic similarity is the top 10 most common words of each
topic.  Each topic is compared to each other topic in the system,
given a certain threshold of top 10 word similarity, such as 8 out of
top 10 matching words.  10 words were chosen because often people care
about top 10 comparison and it allowed for some common words to exist
but not cause topics to match so closely.  We then apply the
transitive closure on similar topics to this topic similarity matrix;
this is similar to modelling topics as nodes and similarity as arcs,
then fill flooding along the similarity arcs until we have partitioned
the topics into clusters of similar topics. Figure \ref{fig:closure}
illustrates clustering of topics by topic similarity. These clusters
of topics are called trends. A trend is probably interesting if it
contains more than one topic.

@@

When we examined PostgreSQL, we did not find many trends with two or
more topics, using a similarity of $7/10$. $7/10$ was chosen because
it preserved indepedant topics but seemed to be a borderline threshold
value where more serious trends started to appear.

Those trends that we did find
were not very large, lasting only 3 months at most. The first and second
largest trends were dedicated to the work of two external developers:
Dal Zotto, Dan McGuirk. The fifth largest trend related to work by
PostgreSQL developer D'Arcy. Other topics of the larger trends were
changes to the TODO, and time string formating topics
relating to timezones.

@@

We want to explore parameters more across multiple projects, what
motivates the need for more or less topics, can we find a good
estimate of a good choice of number of topics to evaluate. We want to
investigate the effects of overlapping windows on this analysis, do
they make the trends too long? Do they highlight some trends
unnecessarily or are they more because they are time agnostic.

Another avenue of future work is automatic topic labelling. Given a word
distribution we should be able to automatically select a word or term
that describes that distribution. Potential external sources of topic
names include: software engineering taxonomies, ontologies and
standards.
